{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-10-06 thesis/tm.tex
fix retry set fp deadlock issue and finish retry set column
add map basicer mutex tm test

\begin{itemize}
	\item Unit tests and microbenchmarks
	\begin{itemize}
		\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example}/\ref{fig:htm-example-reproduced}.
			%(assertion failure on {\tt count}).
		\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
		\item {\tt counter}: Microbenchmark version of {\tt htm2} which replaces the complex locking failure path with an atomic {\tt xadd}, from \cite{tm-benchmark-cmu}.
		\item {\tt swap}: Microbenchmark that swaps values in an array, from \cite{tm-benchmark-cmu}.
		\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
			% TODO: make sure figure number reference is right
		\item {\tt fig63}: Generalized version of Figure~\ref{fig:retry-set-example},
			contrived to induce as much reduction as possible from the retry sets optimization.
	\end{itemize}
	\item Data structure tests
	\begin{itemize}
		\item {\tt avl\_insert}: AVL tree concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed (spoilers!!).
		\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
		\item {\tt avl\_mutex}: {\tt avl\_fixed} with transactional sections simplified by abstraction into a mutex.
		\item {\tt map\_mutex}: {\tt map\_basic} simplified likewise.
	\end{itemize}
	\item Lock abstraction tests
	\begin{itemize}
		\item {\tt lock}(): Checks that multiple threads using a transactional lock
			cannot access the critical section simultaneously.
			% htm_spinlock and htm_mutex
		\item {\tt lock\_fast}(): Checks that a transactional lock's fast path will not suffer conflict aborts
			if its client threads' critical sections are independent.
			% htm_spinlock_fastpath and htm_mutex_fastpath
	\end{itemize}
		These are each parameterized over implementations {\tt spinlock} (from \cite{spinlock-rtm-github}),
		{\tt spin\_\allowbreak{}fixed} (spoilers!!),
		and {\tt mutex} (replaces the spinlock with a Landslide-annotated P2 mutex to reduce state space size).
\end{itemize}

@@

			& 3,2 &\ETAdag{1y 299d}	&\ETAdag{4510472}& \ETAdag{qqq}	& \ETAdag{qqq}	&\ETAdag{10d 16h}&\ETAdag{2322150}\\

@@

@@

@@

			& 2,2 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	&\ETAdag{36393y}&\ETAdag{35094477}\\

@@

			& 3,1 & \ETAdag{5y 292d}&\ETAdag{2873664642}&\ETAdag{qqq}& \ETAdag{qqq}	&\ETAdag{97y 100d}&\ETAdag{1698185036} \\

@@

@@

			& 2,2 & \ETAdag{1y 176d}&\ETAdag{75516602}&\ETAdag{qqq}	& \ETAdag{qqq}	&\ETAdag{13d 11h}&\ETAdag{565334} \\

@@

@@

			& 3,1 &\ETAdag{6y 133d}&\ETAdag{1847957714}&\ETAdag{qqq}& \ETAdag{qqq}	&\ETAdag{4y 156d}&\ETAdag{888242178} \\

@@

@@

\begin{table}[t]
	\begin{center}
		\footnotesize
		\begin{tabular}{cc}
		\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf STM ({\tt -A -S})} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf K,N  & \em (or \ETAdag{\bf \em est.})
			                      & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.41} & \ints{4} \\
			({\tt spinlock})
			& 2,2 & \cpu{198.77} & \ints{1702} \\
			% if lock(spin_fixed) will time out on 2,3 and 3,2, this is guaranteed to aswell
			& 3,1 & \cpu{35.28} & \ints{246} \\
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.55} & \ints{4} \\
			({\tt spin\_fixed})
			& 2,2 & \cpu{105.57} & \ints{998} \\
			& 2,3 & \ETAdag{33h 13m} & \ETAdag{321553} \\
			% [JOB 4] progress: 138911/321553 brs (43.200015%), ETA 23h 12m 50s (elapsed 9h 59m 48s)
			% total CPU time consumed: 10h 4s (36004577784 usecs) (core saturation: 12%)
			& 3,1 & \cpu{28.27} & \ints{186} \\
			& 3,2 & \ETAdag{13y 281d} & \ETAdag{1443676} \\
			% [JOB 4] progress: 169893/1443676 brs (11.768080%), ETA 13y 280d 17h 30m 54s (elapsed 9h 59m 48s)
			& 4,1 & \ETAdag{16h 26m} & \ETAdag{432628} \\ % was: 30408, 282084 pre mkrun
			% [JOB 4] progress: 286910/432628 brs (66.317896%), ETA 6h 26m 25s (elapsed 9h 59m 48s)
			% total CPU time consumed: 10h 4s (36004288982 usecs) (core saturation: 12%)
			\hline
			{\tt lock}
			& 2,1 & \cpu{3.44} & \ints{4}		\\
			({\tt mutex})
			& 2,2 & \cpu{16.83} & \ints{180}	\\
			& 2,3 & \cpu{405.21} & \ints{9372}	\\
			& 2,4 & \cpu{24999.68} & \ints{489480}	\\ % was: 15266 sec 384512 pre mkrun
			% total CPU time consumed: 6h 56m 39s (24999680839 usecs) (core saturation: 12%)
			& 3,1 & \cpu{15.00} & \ints{132}	\\
			%& 3,2 & \cpu{29432.87} & \ints{707086}	\\ % pre mkrun fix (let alone bottom1)
			& 3,2 & \ETAdag{26h 42m} & \ETAdag{1223955} \\
			% [JOB 2] progress: 714882/1223955 brs (58.407517%), ETA 16h 41m 58s (elapsed 9h 59m 55s)
			% total CPU time consumed: 10h 3s (36003621109 usecs) (core saturation: 12%)
			& 4,1 & \cpu{665.89} & \ints{15064}	\\
			\hline
			{\tt lock\_fast}
			& 2,1 & \cpu{3.25} & \ints{1}	\\
			({\tt spin\_fixed})
			& 9,9 & \cpu{4.61} & \ints{1}	\\
			\hline
			{\tt lock\_fast}
			& 2,1 & \cpu{3.19} & \ints{1}	\\
			({\tt mutex})
			& 9,9 & \cpu{4.62} & \ints{1}	\\
			\end{tabular}
			&
			\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf Non-transactional} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf K,N & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			\hline
			\hline
			{\tt avl\_mutex}
			& 2,1 & \cpu{3.48} & \ints{7} \\
			& 2,2 & \cpu{6.25} & \ints{85} \\
			& 2,3 & \cpu{24.69} & \ints{561} \\
			& 2,4 & \cpu{217.98} & \ints{4984} \\
			& 3,1 & \cpu{8.26} & \ints{129} \\
			& 3,2 & \cpu{1403.46} & \ints{30653} \\
			& 3,3 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 4,1 & \cpu{199.96} & \ints{4488} \\
			& 4,2 & \ETAdag{qqq} & \ETAdag{qqq} \\
			\cline{1-4}
			{\tt map\_mutex}
			& 2,1 & \cpu{39.81} & \ints{83} \\
			& 2,2 & \ETAdag{26h 21m} & \ETAdag{1085126} \\
			% [JOB 9] progress: 496530/1085126 brs (45.757812%), ETA 16h 21m 0s (elapsed 9h 59m 38s)
			% total CPU time consumed: 10h 19s (36019507965 usecs) (core saturation: 12%)
			& 3,1 & \ETAdag{173d 17h} & \ETAdag{12572187} \\
			% [JOB 9] progress: 496281/12572187 brs (3.947451%), ETA 173d 6h 38m 55s (elapsed 9h 59m 43s)
			% total CPU time consumed: 10h 20s (36020148608 usecs) (core saturation: 12%)
			\\
			\\
			\\
			\\
			\\
		\end{tabular}
		\\
		\\
		(a) Verifying HTM locks alone.
		&
		(b) Verifying the lock's client code.
		\end{tabular}
	\end{center}
	\caption{Continuation of Table~\ref{tab:tm-verifs},
		demonstrating abstraction reduction \cite{dbug-phdthesis}
		on the {\tt avl\_fixed} and {\tt map\_basic} tests
		by verifying HTM mutex implementations separately.
		Tested with STM semantics, as both lock implementations include a retry loop.}
	\label{tab:tm-verifs2}
\end{table}

>>>>

\begin{itemize}
	\item Unit tests and microbenchmarks
	\begin{itemize}
		\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example}/\ref{fig:htm-example-reproduced}.
			%(assertion failure on {\tt count}).
		\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
		\item {\tt counter}: Microbenchmark version of {\tt htm2} which replaces the complex locking failure path with an atomic {\tt xadd}, from \cite{tm-benchmark-cmu}.
		\item {\tt swap}: Microbenchmark that swaps values in an array, from \cite{tm-benchmark-cmu}.
		\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
			% TODO: make sure figure number reference is right
		\item {\tt fig63}: Generalized version of Figure~\ref{fig:retry-set-example},
			contrived to induce as much reduction as possible from the retry sets optimization.
	\end{itemize}
	\item Data structure tests
	\begin{itemize}
		\item {\tt avl\_insert}: AVL tree concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed (spoilers!!).
		\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test \cite{tm-benchmark-cmu}.
		\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
		\item {\tt avl\_mutex}: {\tt avl\_fixed} with transactional sections simplified by abstraction into a mutex.
		\item {\tt map\_mutex}: {\tt map\_basic} simplified likewise.
		\item {\tt map\_basicer\_mutex}: {\tt map\_basicer} simplified likewise.
	\end{itemize}
	\item Lock abstraction tests
	\begin{itemize}
		\item {\tt lock}(): Checks that multiple threads using a transactional lock
			cannot access the critical section simultaneously.
			% htm_spinlock and htm_mutex
		\item {\tt lock\_fast}(): Checks that a transactional lock's fast path will not suffer conflict aborts
			if its client threads' critical sections are independent.
			% htm_spinlock_fastpath and htm_mutex_fastpath
	\end{itemize}
		These are each parameterized over implementations {\tt spinlock} (from \cite{spinlock-rtm-github}),
		{\tt spin\_\allowbreak{}fixed} (spoilers!!),
		and {\tt mutex} (replaces the spinlock with a Landslide-annotated P2 mutex to reduce state space size).
\end{itemize}

@@

			& 3,2 &\ETAdag{1y 299d}&\ETAdag{4510472}&\ETAdag{294d 5h}&\ETAdag{8185793}&\ETAdag{10d 16h}&\ETAdag{2322150}\\
			% argh.. since re running the retry set i feel like i have to re run the others...
			% the small state spaces though all end up similar perf numbers though;
			% i wonder why this one alone broke 1mil?

@@

			% [JOB 4] progress: 1133447/8185793 brs (13.846513%), ETA 293d 19h 13m 36s (elapsed 9h 59m 55s) % good retry

@@

			% old, owned retry set --
			% ls-id-log-20180817141549.log.ByMMKx.gz
			% [JOB 4] BUG FOUND: landslide-trace-1534529755.156051.html (981774 interleavings; 132 preemptions; ...)
			% it's almost certainly getting retry set blocked
			% so you just have to detect a retry set block deadlock, and un deadlock it
			% this should not affect other tests (not even in cpu time (beyond icache considerations))
			% when you fix it and re run it, remember to use leave-logs, and put a ALWAYS prince in landslide
			% to see when it happens

@@

			& 2,2 &\ETAdag{58880y}&\ETAdag{4.62$\times\mathit{10^7}$}&\ETAdag{51191y}&\ETAdag{3.33$\times\mathit{10^7}$}&\ETAdag{36393y}&\ETAdag{3.51$\times\mathit{10^7}$}\\
			% [JOB 19] progress: 366191/46150466 brs (0.793472%), ETA 58879y 219d 6h 4m 16s (elapsed 9h 59m 28s)
			% [JOB 19] progress: 351391/33318009 brs (1.054658%), ETA 51191y 111d 17h 37m 55s (elapsed 9h 59m 28s)

@@

			& 3,1 & \ETAdag{5y 292d}&\ETAdag{2.87$\times\mathit{10^9}$}&\ETAdag{735y 212d}&\ETAdag{3.35$\times\mathit{10^8}$}&\ETAdag{97y 100d}&\ETAdag{1.70$\times\mathit{10^9}$} \\

@@

			% [JOB 25] progress: 18151/335288683 brs (0.005414%), ETA 735y 211d 9h 1m 8s (elapsed 22m 25s)

@@

			& 2,2 & \ETAdag{1y 176d}&\ETAdag{7.55$\times\mathit{10^7}$}&\ETAdag{11y 94d}&\ETAdag{9.57$\times\mathit{10^7}$}&\ETAdag{13d 11h}&\ETAdag{565334} \\

@@

			% [JOB 22] progress: 128715/95718464 brs (0.134472%), ETA 11y 93d 8h 25m 39s (elapsed 4h 52m 39s)

@@

			& 3,1 &\ETAdag{6y 133d}&\ETAdag{1.85$\times\mathit{10^9}$}&\ETAdag{8y 282d}&\ETAdag{7.90$\times\mathit{10^8}$}&\ETAdag{4y 156d}&\ETAdag{8.88$\times\mathit{10^8}$} \\

@@

			% [JOB 22] progress: 282358/789973638 brs (0.035743%), ETA 8y 281d 14h 50m 3s (elapsed 9h 49m 54s)

@@

\begin{table}[t]
		\begin{tabular}{cc}
		\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf STM ({\tt -A -S})} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf K,N  & \em (or \ETAdag{\bf \em est.})
			                      & \em (or \ETAdag{\bf \em est.}) \\
			{\tt lock}
			& 2,1 & \cpu{3.41} & \ints{4} \\
			({\tt spinlock})
			& 2,2 & \cpu{198.77} & \ints{1702} \\
			% if lock(spin_fixed) will time out on 2,3 and 3,2, this is guaranteed to aswell
			& 3,1 & \cpu{35.28} & \ints{246} \\
			{\tt lock}
			& 2,1 & \cpu{3.55} & \ints{4} \\
			({\tt spin\_fixed})
			& 2,2 & \cpu{105.57} & \ints{998} \\
			& 2,3 & \ETAdag{33h 13m} & \ETAdag{321553} \\
			% [JOB 4] progress: 138911/321553 brs (43.200015%), ETA 23h 12m 50s (elapsed 9h 59m 48s)
			% total CPU time consumed: 10h 4s (36004577784 usecs) (core saturation: 12%)
			& 3,1 & \cpu{28.27} & \ints{186} \\
			& 3,2 & \ETAdag{13y 281d} & \ETAdag{1443676} \\
			% [JOB 4] progress: 169893/1443676 brs (11.768080%), ETA 13y 280d 17h 30m 54s (elapsed 9h 59m 48s)
			& 4,1 & \ETAdag{16h 26m} & \ETAdag{432628} \\ % was: 30408, 282084 pre mkrun
			% [JOB 4] progress: 286910/432628 brs (66.317896%), ETA 6h 26m 25s (elapsed 9h 59m 48s)
			% total CPU time consumed: 10h 4s (36004288982 usecs) (core saturation: 12%)
			{\tt lock}
			& 2,1 & \cpu{3.44} & \ints{4}		\\
			({\tt mutex})
			& 2,2 & \cpu{16.83} & \ints{180}	\\
			& 2,3 & \cpu{405.21} & \ints{9372}	\\
			& 2,4 & \cpu{24999.68} & \ints{489480}	\\ % was: 15266 sec 384512 pre mkrun
			% total CPU time consumed: 6h 56m 39s (24999680839 usecs) (core saturation: 12%)
			& 3,1 & \cpu{15.00} & \ints{132}	\\
			%& 3,2 & \cpu{29432.87} & \ints{707086}	\\ % pre mkrun fix (let alone bottom1)
			& 3,2 & \ETAdag{26h 42m} & \ETAdag{1223955} \\
			% [JOB 2] progress: 714882/1223955 brs (58.407517%), ETA 16h 41m 58s (elapsed 9h 59m 55s)
			% total CPU time consumed: 10h 3s (36003621109 usecs) (core saturation: 12%)
			& 4,1 & \cpu{665.89} & \ints{15064}	\\
			& 2,1 & \cpu{3.25} & \ints{1}	\\
			({\tt spin\_fixed})
			& 9,9 & \cpu{4.61} & \ints{1}	\\
			& 2,1 & \cpu{3.19} & \ints{1}	\\
			({\tt mutex})
			& 9,9 & \cpu{4.62} & \ints{1}	\\
			&
			\begin{tabular}{cc||r|r}
			& & \multicolumn{2}{c}{\bf Non-transactional} \\
			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
			\bf test & \bf K,N & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
			{\tt avl\_mutex}
			& 2,1 & \cpu{3.48} & \ints{7} \\
			& 2,2 & \cpu{6.25} & \ints{85} \\
			& 2,3 & \cpu{24.69} & \ints{561} \\
			& 2,4 & \cpu{217.98} & \ints{4984} \\
			& 3,1 & \cpu{8.26} & \ints{129} \\
			& 3,2 & \cpu{1403.46} & \ints{30653} \\
			& 3,3 & \ETAdag{qqq} & \ETAdag{qqq} \\
			& 4,1 & \cpu{199.96} & \ints{4488} \\
			& 4,2 & \ETAdag{qqq} & \ETAdag{qqq} \\
			\cline{1-4}
			{\tt map\_mutex}
			& 2,1 & \cpu{39.81} & \ints{83} \\
			& 2,2 & \ETAdag{26h 21m} & \ETAdag{1085126} \\
			% [JOB 9] progress: 496530/1085126 brs (45.757812%), ETA 16h 21m 0s (elapsed 9h 59m 38s)
			% total CPU time consumed: 10h 19s (36019507965 usecs) (core saturation: 12%)
			& 3,1 & \ETAdag{173d 17h} & \ETAdag{12572187} \\
			% [JOB 9] progress: 496281/12572187 brs (3.947451%), ETA 173d 6h 38m 55s (elapsed 9h 59m 43s)
			% total CPU time consumed: 10h 20s (36020148608 usecs) (core saturation: 12%)
			\cline{1-4}
			& 2,1 & \cpu{14.34} & \ints{9} \\
			{\tt \_mutex}
			& 2,2 & \cpu{147.27} & \ints{2953} \\
			& 2,3 & \cpu{12946.21} & \ints{244691} \\
			& 3,1 & \cpu{637.81} & \ints{12707} \\
			& 3,2 & \cpu{qqq} & \ints{qqq} \\
			& 4,1 & \ETAdag{33h 33m} & \ETAdag{1778661} \\
			% [JOB 3] progress: 601402/1778661 brs (33.812059%), ETA 23h 33m 16s (elapsed 9h 59m 55s)
			\\
		\\
		\\
		(a) Verifying HTM locks alone.
		&
		(b) Verifying the lock's client code.
	\caption{Continuation of Table~\ref{tab:tm-verifs},
		demonstrating abstraction reduction \cite{dbug-phdthesis}
		on the {\tt avl\_fixed} and {\tt map\_basic} tests
		by verifying HTM mutex implementations separately.
		Tested with STM semantics, as both lock implementations include a retry loop.}
	\label{tab:tm-verifs2}
